### 2014-10-22 thesis/method/classification.tex
Added outline of decision trees

\section{C4.5 Decision Tree (DT)}

>>>>

\section{C4.5 Decision Tree (DT)}
Decision trees (DT) can be specified recursively and are an example of a
``divide-and-conquer'' approach to learning. First, we choose a feature to
place at the root node and then create one branch for each possible value.
This partitions the data set into subsets, one for every value that the
feature takes. We repeat this for each branch, choosing different features
each time and continuing to divide the data set into progressively smaller
subsets, growing the tree.
If all examples in a subset have the same class at any point during
this process, we stop splitting that subset.

The only thing left to specify
now is how to decide which attribute to split on. Note that when splitting,
we would like to create subsets that are as ``pure''
-- that is, containing as many
instances of the same class -- as possible, so that the splitting may
terminate earlier. Since we would like our trees to be as small as possible to
prevent overfitting and allow the best generalisation ability,
we would therefore like to choose the feature to split on such that the
created subsets were as pure as possible. The measure of purity used in
learning decision trees is the \textit{information}, and is measured in
\textit{bits}. At any node (or subset) of the tree, this measure represents
the expected amount of information that would be needed to classify a new
example, given that the example reached the node. The information at each
node is calculated using the \textit{entropy} function:
\begin{equation*}
\mathrm{entropy}(p_1,p_2,\ldots,p_n) = -p_1\mathrm{log}p_1 -p_2\mathrm{log}p_2 \ldots -p_n\mathrm{log}p_n
\end{equation*}
Given $k$ classes, $c_1,\ldots,c_k$ is the number of instances in each
class at a node, and $\sum_{i=1}^k c_i = c$ is the total number of instances
at that node, the information of that node is then:
\begin{equation*}
\mathrm{info}(c_1,\ldots,c_k) = \mathrm{entropy}\left(\frac{c_1}{c},\frac{c_2}{c},\ldots,\frac{c_k}{c}\right)
\end{equation*}

To find the attribute to split on, we assume we are splitting on a particular
attribute, and calculate the information at the node given the split. We do
this for each attribute, and compute the \textit{information gain} from
splitting on that attribute:
\begin{equation*}
\mathrm{gain}(feature) = \mathrm{info}(\text{child node}) - \mathrm{info}(\text{parent node})
\end{equation*}
The feature that is chosen is therefore the one with the highest information
gain.

Decision trees are fast to train and produce models that are straightforward
for humans to interpret, the latter of which is particularly valuable in a
clinical environment. We will use the C4.5 tree induction algorithm, which is
considered the state-of-the-art and handles different types of feature values
as well as missing values.
